---
title: "Getting Islands from EEZs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(sf)
  library(janitor)
  library(here)
  library(tidyverse)
})
```


First load the EEX v10 layer, select only Kiribati

```{r}
EEZ <- read_sf(dsn = here("raw_data","spatial","World_EEZ_v10_20180221"),
               layer = "eez_v10",
               quiet = T,
               stringsAsFactors = F) %>% 
  clean_names( )%>% 
  filter(iso_ter1 %in% c("KIR"))
```

Keep only the first two polygons for visualization purposes

```{r}
EEZ_kir <- EEZ %>% 
  head(2)
```

Look at the dimension, it has two polygons (EEZ) with 23 variables

```{r}
dim(EEZ_kir)
```

Visualize it

```{r}
ggplot(EEZ_kir) +
  geom_sf() +
  theme_bw()
```

To extract the coastline, convert `EEZ_kir` to boundaries and then to linestring, and then again to polygon

```{r}
all_features <- EEZ_kir %>% 
  st_boundary() %>% 
  st_cast(to = "LINESTRING") %>% 
  st_cast(to = "POLYGON") %>% 
  mutate(id = 1:length(iso_ter1))
```

Check dimensions, now we have 37 polygons, one for each island and one for each EEZ blob

```{r}
dim(all_features)
```


We don't want the EEZs for this. We plot and identify which ones these are. Some facets look like there's nothing, but those are just very small islands.

```{r}
ggplot(all_features) + 
  geom_sf() +
  facet_wrap(~id) +
  theme_bw()
```

We see that id's 1 and 12 are EEZs, and we sant to filter them out. Here I filter them by the id I created, but perhaps area could be a better indicator. (remove all features with area more than X km^2)

```{r}
coastline <- all_features %>% 
  filter(!id %in% c(1, 12))
```

Check dimensions again
```{r}
dim(coastline)
```

And plot the island coastline

```{r}
ggplot(coastline) +
  geom_sf() +
  theme_bw()
```

The entire process is:

```{r}
coastline <- read_sf(dsn = here("raw_data","spatial","World_EEZ_v10_20180221"),
               layer = "eez_v10",
               quiet = T,
               stringsAsFactors = F) %>% 
  clean_names( )%>% 
  filter(iso_ter1 %in% c("KIR")) %>% 
  head(2) %>% 
  st_boundary() %>% 
  st_cast(to = "LINESTRING") %>% 
  st_cast(to = "POLYGON") %>% 
  mutate(id = 1:length(iso_ter1)) %>% 
  filter(!id %in% c(1, 12)) #a smarter filter can be used here

ggplot(coastline) + 
  geom_sf() +
  theme_bw()
```




