---
title: "latin_america"
author: "Juan Carlos Villaseñor-Derbez"
date: "Last update `r Sys.Date()`"
output: html_document
---

# Ideal outline:

- EEZ-ecoregions
  - Create a shapefile of ecoregions broken down by EEZ `st_intersection()`
  - Calculate the area of these new pieces `st_area()`

- MPAs
  - For polygons
    - Break every MULTIPOLYGON into single POLYGONS `st_cast()`
    - Calculate the area of each of these new polygons
  - For points
    - Create a buffer arround the point, with a radius that creates an equivalent area
    - Convert buffer into a polygon
    - Make sure area of polygon is the same as area in the data
    - Make sure to have an object of class POLYGON and not MULTIPOLYGON
  - Put polygons and points (now also polygons) together
  - Break these new polygons by the eez-ecoregion polygons calculated before
  - Group_by eez-ecoregion and calculate total and relative protection
  
- Fishing effort
  - Mask fishing effort points to keep things within each eez-ecoregion
  
- Put these together
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      warning = F,
                      mesage = F)
```

# Housekeeping

```{r}
## Clear the environment
rm(list = ls())

# Load packages
suppressPackageStartupMessages({
  library(janitor)
  library(here)
  library(lwgeom)
  library(sf)
  library(tidyverse)
})

# Source functions
source(here("scripts","sfc_as_cols.R"))

deg2rad <- function(deg){
  (deg * pi) / (180)
  }
```

```{r}
proj <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

countries <- c("MEX", "GTM", "SLV", "HND", "NIC", "CRI", "PAN", "COL","PER", "ECU", "CHL")
```

```{r}
coastline <- st_read(dsn = here("data", "spatial"),
                     layer = "world_coastline",
                     quiet = T,
                     stringsAsFactors = F) %>% 
  filter(iso3 %in% countries) %>% 
  mutate(coastline = 1) %>% 
  group_by(coastline) %>% 
  summarize(n = n())
```

# Part1 EEZ Ecoregions

## Create a shapefile of ecoregions broken down by EEZ `st_intersection()`

Load EEZ data, filter by countries we want, reproject and keep only relevant variables

```{r}
EEZ <- read_sf(dsn = here("raw_data","spatial","World_EEZ_v10_20180221"),
               layer = "eez_v10",
               quiet = T,
               stringsAsFactors = F) %>% 
  clean_names() %>% 
  filter(iso_ter1 %in% countries) %>% 
  st_simplify(dTolerance = 0.1)%>% 
  st_transform(crs = proj) %>% 
  select(iso3 = iso_ter1)
```

Load ecoregions data, fix names, and reproject

```{r}
meow <- read_sf(dsn = here("raw_data","spatial","MEOW"),
                layer = "meow_ecos",
                quiet = T,
                stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  st_transform(crs = proj)
```

Break ecoregions and EEZ by areas where they intersect, calculate the area and create a code name for each new polygon

```{r}
eez_ecoregion <- EEZ %>% 
  st_intersection(meow) %>% 
  mutate(area = st_area(.),
         eez_ecoregion_area = as.numeric(area) / 1e6,
         eez_ecoregion = paste(iso3, ecoregion, sep = "_")) %>% 
  select(iso3, eez_ecoregion, realm, province, ecoregion, eez_ecoregion_area)
```

Visualize the new parts

```{r}
ggplot() +
  geom_sf(data = coastline, color = "black", fill = "transparent") +
  geom_sf(data = eez_ecoregion, aes(fill = iso3)) +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 8))
```

# Part 2 MPAS

## For polygons

Break every MULTIPOLYGON into single POLYGONS `st_cast()` and calculate the area of each of these new polygons. As soon as we read, we filter out undesired countries and iucn categories. We validate all polygons and proceed to `cast` each MPA into individual polygons. Then we calculate the individual area of eachnew polygon and reproject. We finilize by selecting the variables we care about.

```{r}
mpa_poly <- read_sf(dsn = here("raw_data","spatial","WDPA_Mar2018"),
                    layer = "WDPA_Mar2018_marine-shapefile-polygons",
                    quiet = T,
                    stringsAsFactors = F) %>% 
  janitor::clean_names() %>%
  filter(iso3 %in% countries) %>%
  st_make_valid() %>% 
  st_cast(to = "POLYGON") %>%
  mutate(mpa_area = st_area(.),
         mpa_area = as.numeric(mpa_area) / 1e6) %>%
  st_transform(crs = proj) %>% 
  select(wdpaid, name, iucn_cat, mpa_area) %>% 
  mutate(source = "polygons")
```

## For points

Create a buffer arround the point, with a radius that creates an equivalent area
Convert buffer into a polygon
Make sure area of polygon is the same as area in the data
Make sure to have an object of class POLYGON and not MULTIPOLYGON

```{r}
mpa_points <- read_sf(dsn = here("raw_data", "spatial", "WDPA_Mar2018"),
                      layer = "WDPA_Mar2018_marine-shapefile-points",
                      quiet = T,
                      stringsAsFactors = F) %>% 
  janitor::clean_names() %>%
  filter(iso3 %in% countries) %>% 
  select(wdpaid, name, iucn_cat, rep_m_area, rep_area) %>% 
  sfc_as_cols(names = c("lon", "lat")) %>% 
  mutate(radius = (sqrt(rep_area / pi)) / (cos(deg2rad(lat)) * 111.32)) %>% 
  st_buffer(dist = .$radius) %>% 
  mutate(test_area = st_area(.),
         test_area = as.numeric(test_area) / 1e6,
         diff = test_area - rep_area)

ggplot(mpa_points) +
  geom_density(aes(x = rep_area), fill = "steelblue", alpha = 0.3) +
  geom_density(aes(x = test_area), fill = "green", alpha = 0.3)

ggplot(mpa_points, aes(x = rep_area, y = test_area)) +
  geom_point() +
  geom_abline()

mpa_points <- mpa_points %>% 
  st_transform(crs = proj) %>% 
  select(wdpaid, name, iucn_cat, mpa_area = rep_area) %>% 
  mutate(source = "points")
  
```

Put polygons and points (now also polygons) together
Break these new polygons by the eez-ecoregion polygons calculated before
Group_by eez-ecoregion and calculate total and relative protection

```{r}
eez_ecoregion_protection <- rbind(mpa_poly, mpa_points) %>% 
  st_intersection(eez_ecoregion) %>% 
  mutate(cropped_mpa_area = st_area(.),
         cropped_mpa_area = as.numeric(cropped_mpa_area / 1e6)) %>% 
  group_by(eez_ecoregion, eez_ecoregion_area) %>% 
  summarize(total_protected = sum(cropped_mpa_area)) %>% 
  ungroup() %>% 
  mutate(percent_protected = total_protected / eez_ecoregion_area) %>% 
  st_set_geometry(NULL) %>% 
  select(eez_ecoregion, total_protected, percent_protected)

temp <- eez_ecoregion %>% 
  left_join(eez_ecoregion_protection, by = "eez_ecoregion")

ggplot() +
  geom_sf(data = coastline, color = "black") +
  geom_sf(data = temp, aes(fill = percent_protected)) +
  scale_fill_viridis_c() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 8))
```

# Part 3 Fishing effort

First I read in the fishing effort

```{r}
mean_binned_fishing <- read.csv(here("data", "mean_yearly_fishing_effort.csv")) %>% 
  filter(mean_fishing_hours > 0)
```

Mask fishing effort points to keep things within each eez-ecoregion

```{r}
masked_mean_binned_fishing <- mean_binned_fishing %>% 
  filter(between(lon_bin, -118, -55),
         between(lat_bin, -67, 33)) %>%
  st_as_sf(coords = c("lon_bin", "lat_bin"), crs = proj) %>% 
  st_intersection(y = eez_ecoregion) %>%
  sfc_as_cols(names = c("lon", "lat")) %>% 
  st_set_geometry(NULL) %>% 
  group_by(eez_ecoregion, eez_ecoregion_area) %>% 
  summarize(hours = sum(mean_fishing_hours)) %>% 
  ungroup() %>% 
  mutate(norm_hours = hours / eez_ecoregion_area)
```

```{r}
temp2 <- eez_ecoregion %>% 
  left_join(masked_mean_binned_fishing, by = "eez_ecoregion")

ggplot() +
  geom_sf(data = coastline, color = "black") +
  geom_sf(data = temp2, aes(fill = norm_hours)) +
  scale_fill_viridis_c() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 8))
```

## Masterplot

```{r}
masterdata <- eez_ecoregion %>% 
  left_join(masked_mean_binned_fishing, by = "eez_ecoregion") %>% 
  left_join(eez_ecoregion_protection, by = "eez_ecoregion") %>% 
  replace_na(replace = list(norm_hours = 0, percent_protected = 0))

ggplot(masterdata, aes(x = percent_protected, y = norm_hours, color = ecoregion, shape = iso3)) +
  geom_point(size = 2) +
  geom_vline(xintercept = 0.1, linetype = "dashed") +
  geom_hline(aes(yintercept = mean(norm_hours, na.rm = T)), linetype = "dashed") +
  scale_shape_manual(values = c(1:11)) +
  theme(legend.position = "none")
```

## Now with america

```{r}
america_masterdata <- readRDS(here("data", "america_masterdata.rds"))
```

```{r}
ggplot() +
  geom_sf(data = coastline, color = "black") +
  geom_sf(data = america_masterdata, aes(fill = percent_protected)) +
  scale_fill_viridis_c() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 8))+
  xlim(-180, -20)
```


```{r}
ggplot() +
  geom_sf(data = coastline, color = "black") +
  geom_sf(data = america_masterdata, aes(fill = norm_hours)) +
  scale_fill_viridis_c() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 8)) +
  xlim(-180, 0)
```


```{r}
america_masterdata %>% 
  filter(norm_hours < 3,
         percent_protected < 1) %>% 
  ggplot(aes(x = percent_protected, y = norm_hours, fill = province)) +
  geom_point(size = 4, shape = 21, color = "black", alpha = 0.5) +
  geom_vline(xintercept = 0.1, linetype = "dashed") +
  geom_vline(xintercept = 0.3, linetype = "dashed") +
  theme_bw() +
  scale_fill_brewer(palette = "Paired", name = "Realm") +
  labs(x = "Percent protected", y = "Fishing (hours / km2)", caption = "Fishing effort and percentage of protected areas by Ecoregion and EEZ")
```







